---
title: "homework3"
format: html
editor: visual
link:
---

```{r}
library(tidyverse)
library(here)
library(gt)
library(janitor)
library(readxl)
library(ggplot2)
```

Problem 1

(a) Data Summarizing

For my data, I am interested in my response variable of the amount of miles ran, and how my mileage is influenced by the explanatory variables of caloric intake and sleep duration. I would like to take the mean of each of these explanatory variables, and then use the mean to determine levels of lower than average or higher than average caloric intake or sleep duration. I would then calculate the mean number of miles ran for each group.

This comparison is informative as it could help me gauge the conditions under which I perform best, and help me further understand how fueling and rest influence my running performance.

(b) Visualization

```{r}
#cleaning data 
personal_data <- read_csv(here("data", "ES193DS_running_data .csv")) # read in the data 
personal_data <- personal_data %>% #cleaning the data 
  clean_names()
```

```{r}
# find the mean for sleep
mean_sleep <- mean(personal_data$sleep_duration_hours, na.rm = TRUE) # calculating the mean for sleep, na.rm = TRUE is being used just in case there are NA values in the data 

# filter the column for  above average sleep 
above_avg_sleep <- personal_data %>% 
  filter(sleep_duration_hours > mean_sleep) # creating a new data set where sleep is above average

# filter the column for  below average sleep
below_avg_sleep <- personal_data %>%
  filter(sleep_duration_hours < mean_sleep) # creating a new data set where sleep is below the average
```

```{r}
# find the mean for caloric intake 
mean_calories <- mean(personal_data$caloric_intake, na.rm = TRUE) # calculating the mena f or sleep, na.rm = TRUE is being used just in case there are NA Values in the data 

# filter for above average caloric intake 
above_avg_caloric_intake <- personal_data %>% 
  filter(caloric_intake > mean_calories) # creating a new data set where caloric intake is above average 

# filter for below average caloric intake 
below_avg_caloric_intake <- personal_data %>%
  filter(caloric_intake < mean_calories) # creating a new data set where caloric intake is below average 

```